Class Paper - Introduction to Bayesian Statistics

### 2019-04-22 Paper_Output.Rmd
Additional formatting of tables, comments in Rmarkdown, and importing model parameters from R scripts

---
title: "Class Paper - Introduction to Bayesian Statistics"
author: "Ben Mainwaring"
date: "3/31/2019"
output: html_document
---

@@

@@

```{r, cache = TRUE, fig.align = 'center', fig.height = 4}
par(mfrow = c(1,2), oma = c(0,0,1.5,0), mar = c(5.1, 4.1, 2.1, 2.1))
hist(preSurvey.merged$Fav_AfD, main = '', xlab = "AfD favourability (pre-election)", ylim = c(0,1000), las = 1)
#axis(side = 2, at = c(0, 200, 400, 600, 800, 1000), labels = c(0, 200, 400, 600, 800, 1000), las = 1)
hist(postSurvey.merged$Fav_AfD, main = '', xlab = "AfD favourability (post-election)", ylim = c(0,1000), las = 1)
mtext("Distribution of AfD Support, German Population", side = 3, outer = TRUE, cex = 1.2, font = 2)
```

@@

```{r, cache = TRUE, fig.align = 'center', fig.height = 4, fig.width = 4}
par(mar = c(5.1, 4.1, 3.6, 2.1))
hist(kreiseData$MigrantBackground_Yes_Pct, main = 'Migrant Background by Constituency',
     xlab = "% of population with migrant background")
```

@@

**can these be moved into multiple columns, as well as prettified?**

    Gamma[migrant.yes] ~ N(0.0, 0.25)
    Gamma[income.high] ~ N(-0.1, 0.25)
    Gamma[educ.high] ~ N(-0.1, 0.25)
    Gamma[unemp] ~ N(0.25, 0.625)
    Gamma[age.60+] ~ N(0.2, 0.5)
    Gamma[AfDshare] ~ N(0.33, 0.68)

    Beta[educ.low] ~ N(1.0, 5.0)
    Beta[educ.med] ~ N(0.5, 5.0)
    Beta[gend.male] ~ N(1.0, 5.0)
    Beta[age] ~ N(0.0, 0.2)
    Beta[age^2^] ~ N(0.0, 0.2)
    Beta[unemp^ ~ N(1.0, 5.0)
    Beta[income] ~ N(-0.1, 1.0)

@@

The pre-election model was computed with three Markov Chains. It was run over 50,000 iterations, using a thinning factor of 25, following a 10,000-iteration burn-in period. The model converges well on most variables. However, the strong autocorrelation between $\beta_{age}$ and $\beta_{age^{2}}$ causes some issues. The difference between naive and time-series standard errors for these two variables is `r round(mean(summary(jags_out.pre_miss)$statistics[c(4:5), 4] / summary(jags_out.pre_miss)$statistics[c(4:5), 3]), 1)`x, compared with `r round(mean(summary(jags_out.pre_miss)$statistics[-c(4:5), 4] / summary(jags_out.pre_miss)$statistics[-c(4:5), 3]), 1)`x for other variables in the model. These can be seen in both the traces (which show very slow change between iterations), and the autocorrelation plots (which show very high dependence, even with thinning at n = 25).

```{r, fig.align = 'center', fig.height = 8, fig.width = 10}
par(mfrow = c(2,2))
plot(jags_out.pre_miss[,4], trace = TRUE, density = FALSE, main = "Trace of B[age]", auto.layout = FALSE)
plot(jags_out.pre_miss[,5], trace = TRUE, density = FALSE, main = "Trace of B[age ^ 2]", auto.layout = FALSE)
autocorr.plot(jags_out.pre_miss[[1]][,4], main = "Autocorrelation of B[age]", sub = "chain 1, thin = 25", auto.layout = FALSE)
autocorr.plot(jags_out.pre_miss[[1]][,5], main = "Autocorrelation of B[age ^ 2]", sub = "chain 1, thin = 25", auto.layout = FALSE)
```

@@

```{r, results = 'asis'}
kable(summ.pre_miss$quantiles[c(8:12),c(1,3,5)],
      caption = "Estimates of Gamma coefficients", digits = 4)
```

@@

<!-- **Do I need to display these tables, given that my ultimate focus is on the post-election model?** -->

<!-- ```{r, results = 'asis'} -->
<!-- #coda.samples stores each markov chain as a separate list - use do.call("rbind"...) to convert them to a single data frame -->
<!-- jags_dfout.pre_miss <-  do.call("rbind", jags_out.pre_miss) -->

<!-- #Calculate the correlation and covariance matrix (correlation for easy display, covariance to feed in to priors of a future model) -->
<!-- #And format nicely -->
<!-- corr.pre_miss <- cor(jags_dfout.pre_miss) -->
<!-- cov.pre_miss <- cov(jags_dfout.pre_miss) -->

<!-- #View correlations -->
<!-- kable(corr.pre_miss[1:7,1:7], caption = "Correlations between Beta parameters in posterior", digits = 2) -->

<!-- kable(corr.pre_miss[8:12,8:12], caption = "Correlations between Beta parameters in posterior", digits = 2) -->

<!-- #to fix table output, try kableExtra -->
<!-- #https://community.rstudio.com/t/kableextra-how-to-create-a-table-with-equal-column-width/4386 -->
<!-- #http://haozhu233.github.io/kableExtra/awesome_table_in_html.html -->

<!-- ``` -->

@@

**Auto-update parameters**The post-election model was calculated with 1 million iterations and a thinning parameter of 125, following a 35,000-iteration burn-in period. Initial diangnostics, run on a much shorter Markov Chain, suggested this was neccessary to handle severe autocorrelation problems. Even with a very long model estimation, autocorrelation was an issue for Gamma parameters, and the Beta parameters related to age, as shown below (autocorrelations plots are arbitrarily taken from the first chain, but all chains show a similar pattern).

```{r, fig.align = 'center', fig.height = 8, fig.width = 10}
par(mfrow = c(4,2))
autocorr.plot(jags_out.post_miss[[1]][,c(4:5)], auto.layout = FALSE)
autocorr.plot(jags_out.post_miss[[1]][,c(8:13)], auto.layout = FALSE)

@@

```{r, fig.align = 'center', fig.height = 4, fig.width = 10}
plot(jags_out.post_miss[,c(1,5)], trace = TRUE, density = FALSE)
```

@@

```{r, fig.align = 'center', fig.height = 4, fig.width = 7}
#visualise posterior vs prior for Gamma[migrant.yes]
#use MCMCtrace function, since it enables the plotting of priors against posteriors - but we first need to draw from the distribution of priors
prior_draws <- rnorm(15000, mean =  summ.pre_miss$statistics[8,1], sd = sqrt(cov.pre_miss[8,8] * 1.5))

MCMCtrace(jags_out.post_miss, params = "G\\[migrant.yes\\]", type = "density", priors = prior_draws,
          main_den = "Prior and Posterior Density for G[migrant.yes]", lty_pr = "dashed", xlim = c(-0.06, 0.06),
          PPO_out = FALSE,
          ISB = FALSE, pdf = FALSE)
```

@@

```{r, fig.align = 'center', fig.height = 5, fig.width = 10}
#visualise all gammas
par(mfrow = c(1,2))
MCMCplot(jags_out.post_miss[,c(8:13)], ref_ovl = TRUE, xlim = c(-.20, .20), main = "Post-Election", sz_ax = 1.5) 
MCMCplot(jags_out.pre_miss[,c(8:13)], ref_ovl = TRUE, xlim = c(-.20, .20), main = "Pre-Election", sz_ax = 1.5)

@@

```{r}
#Merge pre- and post-election output into a matrix
beta.compare <- cbind(summ.post_miss$statistics[c(1:7),c(1:2)],  summ.pre_miss$statistics[c(1:7),c(1:2)])
colnames(beta.compare) <- c("Post-Election Mean", "Post-Election SD", "Pre-Election Mean", "Pre-Election SD")
```

@@

```{r, fig.align = 'center', fig.height = 4, fig.width = 7}
#For each of 8000 iterations * 3 markov chains, generate the predicted value at each age
jags_dfout.post_miss <-  do.call("rbind", jags_out.post_miss)
jags_dfout.pre_miss <-  do.call("rbind", jags_out.pre_miss)
mfx_per_iter.age.post <- apply(jags_dfout.post_miss[,c(4,5)], 
      1, function(x){
          mfx_out <- (age_seq * x[1]) + (age_seq ^ 2 * x[2])
          return(mfx_out)
      })
mfx_per_iter.age.pre <- apply(jags_dfout.pre_miss[,c(4,5)], 
       1, function(x){
           mfx_out <- (age_seq * x[1]) + (age_seq ^ 2 * x[2])
           return(mfx_out)
       })

peak_per_iter.age.post <- apply(mfx_per_iter.age.post, 2, which.max) + 17
peak_per_iter.age.pre <- apply(mfx_per_iter.age.pre, 2, which.max) + 17

peak_density.age.post <- density(peak_per_iter.age.post, bw = 1.5)
peak_density.age.pre <- density(peak_per_iter.age.pre, bw = 1.5)

peak_density.age.pre <- density(peak_per_iter.age.pre, bw = 1.5)
plot(peak_density.age.post, xlim = c(18,40), main = "Distribution of Age of Peak AfD support")
lines(peak_density.age.pre, lty = "dashed", col = "red")
legend(x = "topleft", legend = c("Post-Election", "Pre-Election"),
       lty = c("solid", "dashed"), col = c("black", "red"))

>>>>

---
title: "Class Paper - Introduction to Bayesian Statistics"
author: "Ben Mainwaring"
date: "4/23/2019"
output: html_document
---

@@

coefnames <- c("B[educ.low]", "B[educ.med]", "B[gend.male]", "B[age]", "B[age ^ 2]", "B[unemp]", "B[income]",
                   "G[migrant.yes]", "G[income.high]", "G[educ.high]", "G[unemp]", "G[age.60+]", "G[vote.AfD]")

@@

```{r, cache = TRUE, fig.align = 'center', fig.height = 4}
#Plot histograms of AfD support in both survey wave

par(mfrow = c(1,2), oma = c(0,0,1.5,0), mar = c(5.1, 4.1, 2.1, 2.1))
hist(preSurvey.merged$Fav_AfD, main = '', xlab = "AfD favourability (pre-election)", ylim = c(0,1000), las = 1)
#axis(side = 2, at = c(0, 200, 400, 600, 800, 1000), labels = c(0, 200, 400, 600, 800, 1000), las = 1)
hist(postSurvey.merged$Fav_AfD, main = '', xlab = "AfD favourability (post-election)", ylim = c(0,1000), las = 1)
mtext("Distribution of AfD Support, German Population", side = 3, outer = TRUE, cex = 1.2, font = 2)
```

@@

```{r, cache = TRUE, fig.align = 'center', fig.height = 4, fig.width = 4}
#Plot histograms of migrant background (key IV)

par(mar = c(5.1, 4.1, 3.6, 2.1))
hist(kreiseData$MigrantBackground_Yes_Pct, main = 'Migrant Background by Constituency',
     xlab = "% of population with migrant background")
```

@@

```{r, results = 'asis'}
#Define priors for Beta and Gamma in matrices, and ten output matrices to table

#Define priors
B_priors <- matrix(data = c(c(1.0, 0.5, 1.0, 0.0, 0.0, 1.0, -0.1), 
                            1 /  c(0.2, 0.2, 0.2, 5.0, 5.0, 0.2, 1.0)), nrow = 2, byrow = TRUE)
colnames(B_priors) <- coefnames[1:7]
rownames(B_priors) <- c("Mean", "SD")

G_priors <- matrix(data = c(c(0.0, -0.1, -0.1, 0.25, 0.2, 0.33), (abs(c(0.2, -0.1, -0.1, 0.25, 0.2, 0.33) * 2.5) ^ 2)),
                   nrow = 2, byrow = TRUE)
colnames(G_priors) <- coefnames[7:12]
rownames(G_priors) <- c("Mean", "SD")

#Output tables via kable function
kable(B_priors, 
      caption = "Priors for Beta (pre-election survey)", digits = 2, format.args = list(nsmall = 2))
kable(G_priors,
      caption = "Priors for Gamma (pre-election survey)", digits = 2, format.args = list(nsmall = 2))
```

@@

The pre-election model was computed with three Markov Chains. It was run over `r format(iterations.pre, big.mark = ",")` iterations, using a thinning factor of `r thin.pre`, following a `r format(burnin.pre, big.mark = ",")`-iteration burn-in period. The model converges well on most variables. However, the strong autocorrelation between $\beta_{age}$ and $\beta_{age^{2}}$ causes some issues. The difference between naive and time-series standard errors for these two variables is `r round(mean(summary(jags_out.pre_miss)$statistics[c(4:5), 4] / summary(jags_out.pre_miss)$statistics[c(4:5), 3]), 1)`x, compared with `r round(mean(summary(jags_out.pre_miss)$statistics[-c(4:5), 4] / summary(jags_out.pre_miss)$statistics[-c(4:5), 3]), 1)`x for other variables in the model. These can be seen in both the traces (which show very slow change between iterations), and the autocorrelation plots (which show very high dependence, even with thinning at n = 25).

```{r, fig.align = 'center', fig.height = 8, fig.width = 10}
#Output trace of coefficients for age

par(mfrow = c(2,2))
plot(jags_out.pre_miss[,4], trace = TRUE, density = FALSE, main = "Trace of B[age]", auto.layout = FALSE)
plot(jags_out.pre_miss[,5], trace = TRUE, density = FALSE, main = "Trace of B[age ^ 2]", auto.layout = FALSE)
autocorr.plot(jags_out.pre_miss[[1]][,4], main = "Autocorrelation of B[age]", sub = "chain 1, thin = 25", auto.layout = FALSE)
autocorr.plot(jags_out.pre_miss[[1]][,5], main = "Autocorrelation of B[age ^ 2]", sub = "chain 1, thin = 25", auto.layout = FALSE)
```

@@

```{r, results = 'asis'}
#Plot table of constituency-level coefficients
kable(summ.pre_miss$quantiles[c(8:12),c(1,3,5)],
      caption = "Estimates of Gamma coefficients", digits = 4)
```

@@

@@

The post-election model was calculated with `r format(iterations.post, big.mark = ",")` iterations and a thinning parameter of `r thin.post`, following a `r format(burnin.post, big.mark = ",")`-iteration burn-in period. Initial diangnostics, run on a much shorter Markov Chain, suggested this was neccessary to handle severe autocorrelation problems. Even with a very long model estimation, autocorrelation was an issue for Gamma parameters, and the Beta parameters related to age, as shown below (autocorrelations plots are arbitrarily taken from the first chain, but all chains show a similar pattern).

```{r, fig.align = 'center', fig.height = 8, fig.width = 10}
#Plot autocorrelations of Betas for age, and all gammas

par(mfrow = c(4,2))
autocorr.plot(jags_out.post_miss[[1]][,c(4:5)], auto.layout = FALSE)
autocorr.plot(jags_out.post_miss[[1]][,c(8:13)], auto.layout = FALSE)

@@

```{r, fig.align = 'center', fig.height = 4, fig.width = 10}
#Plot trace of Beta[age ^ 2] (a problem variable) against Beta[educ.low] (a well-behaved one)
plot(jags_out.post_miss[,c(1,5)], trace = TRUE, density = FALSE)
```

@@

```{r, fig.align = 'center', fig.height = 4, fig.width = 7}
#visualise posterior vs prior for Gamma[migrant.yes]

#use MCMCtrace function, since it enables the plotting of priors against posteriors - but we first need to draw from the distribution of priors
prior_draws <- rnorm(15000, mean =  summ.pre_miss$statistics[8,1], sd = sqrt(cov.pre_miss[8,8] * 1.5))

#Plot
MCMCtrace(jags_out.post_miss, params = "G\\[migrant.yes\\]", type = "density", priors = prior_draws,
          main_den = "Prior and Posterior Density for G[migrant.yes]", lty_pr = "dashed", xlim = c(-0.06, 0.06),
          PPO_out = FALSE,
          ISB = FALSE, pdf = FALSE)
```

@@

```{r, fig.align = 'center', fig.height = 5, fig.width = 10}
#visualise all gammas

par(mfrow = c(1,2))
MCMCplot(jags_out.post_miss[,c(8:13)], ref_ovl = TRUE, xlim = c(-.20, .20), main = "Post-Election", sz_ax = 1.5) 
MCMCplot(jags_out.pre_miss[,c(8:13)], ref_ovl = TRUE, xlim = c(-.20, .20), main = "Pre-Election", sz_ax = 1.5)

@@

```{r}
#Merge pre- and post-election output into a matrix - don't print it yet, but the inline code below depends on it
beta.compare <- cbind(summ.post_miss$statistics[c(1:7),c(1:2)],  summ.pre_miss$statistics[c(1:7),c(1:2)])
colnames(beta.compare) <- c("Post-Election Mean", "Post-Election SD", "Pre-Election Mean", "Pre-Election SD")
```

@@

```{r, fig.align = 'center', fig.height = 4, fig.width = 7}
#Visualise the age which maximises AfD support
#More specifically - chart the distribution of the age that maximises AfD support, across iteerations of the Markov chain

#For each of 8000 iterations * 3 markov chains, generate the predicted value at each age
jags_dfout.post_miss <-  do.call("rbind", jags_out.post_miss)
jags_dfout.pre_miss <-  do.call("rbind", jags_out.pre_miss)
mfx_per_iter.age.post <- apply(jags_dfout.post_miss[,c(4,5)], 
      1, function(x){
          mfx_out <- (age_seq * x[1]) + (age_seq ^ 2 * x[2])
          return(mfx_out)
      })
mfx_per_iter.age.pre <- apply(jags_dfout.pre_miss[,c(4,5)], 
       1, function(x){
           mfx_out <- (age_seq * x[1]) + (age_seq ^ 2 * x[2])
           return(mfx_out)
       })

#Identify the age that maximises AfD support, for each iteration
#NB: since our survey is 18+, this first item in age_seq is "18" - so if which.max returns 1, this corresponds to an age of 18; to calculate age from which.max we add 17
peak_per_iter.age.post <- apply(mfx_per_iter.age.post, 2, which.max) + 17
peak_per_iter.age.pre <- apply(mfx_per_iter.age.pre, 2, which.max) + 17

#Compute and plot kernel densities
peak_density.age.post <- density(peak_per_iter.age.post, bw = 1.5)
peak_density.age.pre <- density(peak_per_iter.age.pre, bw = 1.5)
plot(peak_density.age.post, xlim = c(18,40), main = "Distribution of Age of Peak AfD support")
lines(peak_density.age.pre, lty = "dashed", col = "red")
legend(x = "topleft", legend = c("Post-Election", "Pre-Election"),
       lty = c("solid", "dashed"), col = c("black", "red"))
